### 2025-07-01 Text/Thesis/Chapters/Chapter3.qmd
Update thesis

```{r setup, include=FALSE}
source("setup.R")
```
# PROBLEM STATEMENT AND METHODOLOGY {#sec-Chapter3}

## Bootstrap Correction for  Entropy Estimators

## Hypothesis Testing

## Proposed Test Statistics for Heterogeneity Detection

>>>>

```{r setup, include=FALSE}
source("setup.R")

source("./Code/entropy_gamma_sar.R")
source("./Code/entropy_gI0.R")
source("./Code/gamma_sar_sample.R")
source("./Code/gi0_sample.R")
source("./Code/al_omari_1_estimator.R")
source("./Code/bootstrap_al_omari_1_estimator.R")
source("./Code/renyi_entropy_estimator_v1.R")
source("./Code/bootstrap_renyi_entropy_estimator_v1.R")
source("./Code/entropy_renyi_gamma_sar.R")

source("./Code/vasicek_estimator.R")
source("./Code/van_es_estimator.R")
source("./Code/noughabi_arghami_estimator.R")
source("./Code/correa_estimator.R")
source("./Code/ebrahimi_estimator.R")
source("./Code/bootstrap_correa_estimator.R")

source("./Code/tsallis_entropy_gamma_sar.R")
source("./Code/tsallis_estimator.R")
source("./Code/bootstrap_tsallis_entropy.R")
source("./Code/tsallis_estimator_optimized.R")
source("./Code/bootstrap_tsallis_entropy_optimized.R")

source("./Code/renyi_estimator.R")
source("./Code/bootstrap_renyi_estimator.R")
source("./Code/renyi_entropy_optimized.R")
source("./Code/bootstrap_renyi_estimator_Op.R")

source("./Code/shannon_alomari_estimator.R")
source("./Code/bootstrap_shannon_alomari.R")
source("./Code/shannon_entropy_gamma_sar.R")

source("./Code/functions_sample_bias_mse.R")
source("./Code/functions_sample_bias_mse_1.R")

theme_set(theme_minimal() +
            theme(text = element_text(family = "serif"),
                  legend.position = "bottom"))

```

# METHODOLOGY {#sec-Chapter3}

## SELECTION OF THE BEST SHANNON ENTROPY ESTIMATORS  {#sec:shannon_selection}

Although several nonparametric estimators have been proposed for Shannon entropy, their performance depends on factors such as sample size and the underlying distribution. We evaluate the six estimators described in Section \ref{sec:sh_est}: the Vasicek \eqref{E:Vasi}, Van Es \eqref{E:VanEs}, Ebrahimi et al.\ \eqref{E:Ebrahimi}, Correa \eqref{E:Correa}, Noughabi–Arghami \eqref{E:rohtua}, and Al–Omari \eqref{E:AO} estimators. Their performance is evaluated through a simulation study under fixed conditions. The selection is based on two standard criteria: bias and mean squared error (MSE).

The goal of this analysis is to identify the estimators with the best overall performance, which will then be used in the subsequent stages of the methodology, including bootstrap correction and hypothesis testing. This selection process is exclusive to Shannon entropy, as only a single estimator was considered for Rényi and Tsallis entropies.

Our experimental setup involves the analysis of bias and MSE for each estimator through a Monte Carlo study. For each sample size $n \in \{9, 25, 49, 81, 121\}$, we generate $1000$  independent synthetic samples from $Z \sim \Gamma_{\text{SAR}}(5,1)$. The results are consistent across other values of $\mu$ and $L$.
We adopt the heuristic spacing $m = \left[\sqrt{n} + 0.5\right]$, a choice commonly recommended in the literature [@Wieczorkowski1999].

Figure \ref{fig:Plot_bias_mse-sh} presents the bias and MSE for each of the non-parametric estimators of Shannon entropy. The results are summarized in @tbl-out.

```{=latex}
\begin{figure}[H]
\caption{Performance of Shannon entropy estimators under $\Gamma_{\text{SAR}}(5,1)$.}\label{fig:Plot_bias_mse-sh}
{\centering \includegraphics[width=1\linewidth]{Figures/Plot_bias_shannon_L5b-1}

}
\end{figure}
```

```{r Simulated_data_L51_cached2, echo=FALSE, message=FALSE, cache=TRUE, autodep=TRUE}
set.seed(1234567890, kind = "Mersenne-Twister")

file_name <- "./Data/results_estimator_sh2.Rdata"

if (file.exists(file_name)) {
  load(file_name)
  message("Loaded existing results from results_estimator_sh2.Rdata")
} else {
  # Parámetros del experimento
  sample_sizes <- c(9, 25, 49, 81, 121 )

# Number of replications
R <-500

# Number of bootstrap replications
#B1 <- 1
mu_values <- c(1)
alpha <- -500
L <- 5

estimators <- list(
  "Vasicek" = vasicek_estimator,
  "Van Es" = van_es_estimator,
  "Ebrahimi" = ebrahimi_estimator,
  "Correa" = correa_estimator,
  "Noughabi" = noughabi_arghami_estimator,
  "AlOmari" = al_omari_1_estimator

  # "Correa" = correa_estimator,
  # "Ebrahimi" = ebrahimi_estimator,
  # "Al Omari" = al_omari_1_estimator,
  # "Correa Bootstrap" = bootstrap_correa_estimator,
  # "Ebrahimi Bootstrap" = bootstrap_ebrahimi_estimator,
  # "Al Omari Bootstrap" = bootstrap_al_omari_1_estimator
)

calculate_results_gi0 <- function(sample_sizes, R, B1, mu_values, alpha, L, estimators) {
  results_list <- list()

  for (mu_val in mu_values) {

    results <- calculate_bias_mse_gi0(sample_sizes, R, B1, mu_val, alpha, L, estimators)
    df <- as.data.frame(results)

    results_list[[as.character(mu_val)]] <- df
  }

  return(results_list)
}

results_gi0 <- calculate_results_gi0(sample_sizes, R, B1, mu_values, alpha, L, estimators)

  save(results_gi0, file = file_name)
  message("Simulation completed and results saved.")
}
```

```{r fig-bias_shannon_L5-011, eval=FALSE, echo=FALSE, message=FALSE, warning=FALSE, fig.show="hide",   out.width="95%", fig.pos = 'H',  fig.cap="Bias and MSE of the entropy estimators for the $\\Gamma_{\\text{SAR}}$, with $L=2$.", fig.width=7,fig.height=3}
#fig.show="hide", 
# eval=FALSE,  #

load("./Data/results_estimator_sh2.Rdata")

estimators_to_plot <- c("Vasicek", "Van Es", "Ebrahimi",  "Correa", "Noughabi", "AlOmari" )
  latex_estimator_names <- c("Vasicek" = expression("$\\widehat{italic(H)}_{V}$"),
                             "Van Es" = expression("$\\widehat{italic(H)}_{VE}$"),
                            "Ebrahimi" = expression("$\\widehat{italic(H)}_{E}$"),
                            "Correa" = expression("$\\widehat{italic(H)}_{C}$"),
                            "Noughabi" = expression("$\\widehat{italic(H)}_{N}$"),
                            "AlOmari" = expression("$\\widehat{italic(H)}_{AO}$"))
selected_estimators_latex <- latex_estimator_names[estimators_to_plot]
mu_values <- 1
combined_plot_gi0 <- generate_plot_gi0_esp(results_gi0, mu_values, selected_estimators_latex, ncol = 1, nrow = 1)

print(combined_plot_gi0)
```

\renewcommand{\arraystretch}{1.8} 
```{r Table_estimators_shannon2, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
#| label: tbl-out
#| tbl-cap: 'Bias and MSE of Shannon entropy estimators under $\Gamma_{\text{SAR}}(5,1)$.'

cat("\\setlength{\\tabcolsep}{4pt}\n")
load("./Data/results_estimator_sh2.Rdata")

estimators_to_table <- c("Vasicek", "Van Es", "Ebrahimi", "Correa", "Noughabi", "AlOmari")

reshaped_results <- purrr::map_dfr(
  results_gi0,
  ~ .x %>% dplyr::filter(Estimator %in% estimators_to_table),
  .id = "mu"
) %>%
  tidyr::pivot_wider(names_from = Estimator, values_from = c(Bias, MSE))

for (j in seq_along(reshaped_results)) {
  v <- reshaped_results[[j]]
  if (is.numeric(v)) {
    if (j == 1) {                    # μ  → entero, scriptstyle
      reshaped_results[[j]] <- sprintf("$\\scriptstyle %d$", v)
    } else if (j == 2) {             # n  → entero “normal”
      reshaped_results[[j]] <- sprintf("$%d$", v)
    } else {                         # resto → 3 decimales + phantom
      reshaped_results[[j]] <- ifelse(
        v < 0,
        sprintf("$%.3f$", v),
        sprintf("$\\phantom{-}%.3f$", v)
      )
    }
  }
}

cols_to_bold <- c(5, 6, 8, 11, 12, 14)  # según el orden final de la tabla

bold_math <- function(x) {
  sub("^\\$(.*)\\$$", "\\$\\\\mathbf{\\1}\\$", x)
}

reshaped_results[cols_to_bold] <- lapply(reshaped_results[cols_to_bold], bold_math)

colnames(reshaped_results) <- c(
  "$\\bm{\\mu}$", "$\\bm{n}$",
  "$\\bm{\\widehat{H}_{\\text{V}}}$",  "$\\bm{\\widehat{H}_{\\text{VE}}}$",
  "$\\bm{\\widehat{H}_{\\text{E}}}$",  "$\\bm{\\widehat{H}_{\\text{C}}}$",
  "$\\bm{\\widehat{H}_{\\text{N}}}$", "$\\bm{\\widehat{H}_{\\text{AO}}}$",
  "$\\bm{\\widehat{H}_{\\text{V}}}$",  "$\\bm{\\widehat{H}_{\\text{VE}}}$",
  "$\\bm{\\widehat{H}_{\\text{E}}}$",  "$\\bm{\\widehat{H}_{\\text{C}}}$",
  "$\\bm{\\widehat{H}_{\\text{N}}}$", "$\\bm{\\widehat{H}_{\\text{AO}}}$"
)

# 5. kable ---------------------------------------------------------------------
knitr::kable(
  reshaped_results,
  #caption   = "Bias and MSE of Shannon entropy estimators under $\\Gamma_{\\text{SAR}}(5,1)$.",
  format    = "latex",
  booktabs  = TRUE,
  align     = "crllllllllllll",  # 
  escape    = FALSE,
 # label     = "#tbl-tablesh",
  centering = FALSE,
 # table.envir = "table",
  #position  = "htb",
  linesep   = ""
) %>%
  add_header_above(c(" ", " ", "Bias" = 6, "MSE" = 6)) %>%
  collapse_rows(columns = 1:2, latex_hline = "major", valign = "middle") %>%
  row_spec(0, align = "r") %>%
  kable_styling(latex_options = "scale_down")

```

As shown in the simulation results, the estimators $\widehat{H}_{\text{C}}$, $\widehat{H}_{\text{E}}$, and $\widehat{H}_{\text{AO}}$ show low bias and achieve convergence for sample sizes larger than 81. These estimators exhibit good behavior in terms of bias and MSE across various parameter combinations.

In contrast, estimator $\widehat{H}_{\text{N}}$ displays the lowest MSE across all scenarios. However, the bias remains unchanged and high for sample sizes larger than 25, indicating that the convergence is not very fast. Additionally, we observe that both $\widehat{H}_{\text{V}}$ and $\widehat{H}_{\text{VE}}$ estimators exhibit larger bias and slower convergence compared to their counterparts. Notably, the $\widehat{H}_{\text{V}}$ estimator shows the highest MSE.

Although the estimators $\widehat{H}_{\text{C}}$, $\widehat{H}_{\text{E}}$, and $\widehat{H}_{\text{AO}}$ demonstrate good performance in terms of bias and MSE for sufficiently large samples (e.g., $n \geq 81$), practical scenarios in SAR image processing often involve much smaller sample sizes. In particular, local texture analysis is typically performed on small windows, such as $7 \times 7$, which correspond to $n = 49$. Under these conditions, bias can become significant, and convergence is slower or incomplete.

To address this issue and improve estimation accuracy in small-sample regimes, we introduce a bootstrap-based correction methodology. The next subsection presents the bias-reduction procedure and the definition of the corrected entropy estimator.

## BOOTSTRAP BIAS CORRECTION FOR ENTROPY ESTIMATORS

The nonparametric estimators of entropy considered in this work include the Shannon entropy $\widehat{H}(\bm{Z})$, the Rényi entropy $\widehat{R}_\lambda(\bm{Z})$, and the Tsallis entropy $\widehat{T}_\lambda(\bm{Z})$. We assume that these estimators are inherently biased; that is, their expected value does not coincide with the true entropy, as shown in Equation \eqref{eq:bias_def}:
\begin{equation}
\operatorname{Bias}(\widehat{\theta}) = \mathbb{E}[\widehat{\theta}] - \theta \neq 0,
\label{eq:bias_def}
\end{equation}
where $\widehat{\theta}$ denotes any of the estimators listed above, and $\theta$ is the corresponding true entropy value.

To reduce this bias, we apply a bootstrap correction method. Let $\bm{Z} = (Z_1, Z_2, \dots, Z_n)$ be a random sample of size $n$ from an unknown distribution. Using bootstrap resampling, we generate $B$ new samples $\bm{Z}^{(1)}, \bm{Z}^{(2)}, \dots, \bm{Z}^{(B)}$ by sampling with replacement from $\bm{Z}$. For each bootstrap sample $\bm{Z}^{(b)}$, we compute the corresponding estimator $\widehat{\theta}(\bm{Z}^{(b)})$.

The bias is then estimated using Equation \eqref{eq:bias_estimate}, which calculates the difference between the average of the bootstrap estimates and the original estimate:
\begin{equation}
\widehat{\operatorname{Bias}} = \frac{1}{B} \sum_{b=1}^{B} \widehat{\theta}(\bm{Z}^{(b)}) - \widehat{\theta}(\bm{Z}).
\label{eq:bias_estimate}
\end{equation}
A bias-corrected estimator is then defined by subtracting the estimated bias from the original estimator, as expressed in Equation \eqref{eq:bias_corrected}:
\begin{equation}
\widetilde{\theta} = \widehat{\theta} - \widehat{\operatorname{Bias}}.
\label{eq:bias_corrected}
\end{equation}
Substituting the estimate from Equation \eqref{eq:bias_estimate} into Equation \eqref{eq:bias_corrected} yields the explicit formula for the corrected estimator:
\begin{equation}
\widetilde{\theta} = 2\widehat{\theta}(\bm{Z}) - \frac{1}{B} \sum_{b=1}^{B} \widehat{\theta}(\bm{Z}^{(b)}).
\label{eq:theta_explicit}
\end{equation}
<!-- This expression follows from the classical bias correction formula: -->
<!-- \begin{equation} -->
<!-- \text{Bias-corrected estimate} = \text{Original estimate} - \text{Bias estimate}. -->
<!-- \label{eq:bias_correction_principle} -->
<!-- \end{equation} -->
Thus, $\widetilde{\theta}$ is an approximately unbiased estimator of $\theta$. The effectiveness of the correction depends on the number of bootstrap replicates $B$ and the characteristics of the underlying distribution, but it systematically reduces bias, particularly when working with small sample sizes.

To distinguish the bootstrap-improved estimators from their original forms, we adopt the following notation: $\widetilde{H}_{\text{V}}$, $\widetilde{H}_{\text{VE}}$, $\widetilde{H}_{\text{E}}$, $\widetilde{H}_{\text{C}}$, $\widetilde{H}_{\text{N}}$, and $\widetilde{H}_{\text{AO}}$, respectively. For the other entropy types, we denote the bootstrap-corrected estimators as $\widetilde{R}_\lambda$ for Rényi entropy and $\widetilde{T}_\lambda$ for Tsallis entropy.

In order to assess the effectiveness of the bootstrap technique, we present comparisons of the bias and MSE between each original non-parametric Shannon entropy estimator and its respective bootstrap-enhanced version, using samples from the $\Gamma_{\text{SAR}}(5,1)$, with $B=100$ bootstrap replicates and varying $n$, as shown in Figures \ref{fig:Firstfigure}--\ref{fig:subfig6}.

```{=latex}
\begin{figure}[htb]
  \caption{Comparing Bias and MSE: original vs. bootstrap estimators, with, $\mu=1$ and $L=5$.}
  \label{fig:all_estimator}
  \centering
  \begin{subfigure}[t]{0.48\textwidth}
    \centering
    \includegraphics[width=\textwidth]{Figures/Plot_bias_mse_vasicek-1}
    \caption{}
    \label{fig:Firstfigure}
  \end{subfigure}
  \hfill
  \begin{subfigure}[t]{0.48\textwidth}
    \centering
    \includegraphics[width=\textwidth]{Figures/Plot_bias_mse_VanEs-1}
    \caption{}
    \label{fig:Secondfigure}
  \end{subfigure}
	\begin{subfigure}[t]{0.48\textwidth}
    \centering
    \includegraphics[width=\textwidth]{Figures/Plot_bias_mse_Ebrahimi-1}
    \caption{}
    \label{fig:subfig3}
  \end{subfigure}
  \hfill
  \begin{subfigure}[t]{0.48\textwidth}
    \centering
    \includegraphics[width=\textwidth]{Figures/Plot_bias_mse_correa-1}
    \caption{}
    \label{fig:subfig4}
  \end{subfigure}
	\begin{subfigure}[t]{0.48\textwidth}
    \centering
    \includegraphics[width=\textwidth]{Figures/Plot_bias_mse_noughabi-1}
    \caption{}
    \label{fig:subfig5}
  \end{subfigure}
  \hfill
  \begin{subfigure}[t]{0.48\textwidth}
    \centering
    \includegraphics[width=\textwidth]{Figures/Plot_bias_mse_AO-1}
    \caption{}
    \label{fig:subfig6}
  \end{subfigure}
\end{figure}
```

The estimators $\widetilde{H}_{\text{V}}$ and $\widetilde{H}_{\text{VE}}$ show consistent bias reduction, though convergence remains slow. For $\widetilde{H}_{\text{E}}$, the correction improves bias only for small $n$, while its MSE steadily decreases. In contrast, the bootstrap technique did not improve the $\widetilde{H}_{\text{N}}$ estimator. This might occur because the original estimator overestimates the entropy values, showing a positive bias. This tendency to overestimate persists with the use of bootstrap, contributing to an increase in bias and MSE.

The most notable improvements are observed with $\widetilde{H}_{\text{C}}$ and $\widetilde{H}_{\text{AO}}$, which achieve both low bias and MSE even for small sample sizes, demonstrating faster convergence.

While the bias-corrected estimator $\widetilde{\theta}$ is theoretically motivated to reduce bias, its effectiveness is not guaranteed for all entropy estimators. In practice, the performance of the bootstrap correction varies depending on the characteristics of the original estimator and the sample size.

To finalize the selection of the non-parametric Shannon entropy estimator used in the following simulations, we consider the execution time of the bootstrap versions of $\widetilde{H}_{\text{C}}$ and $\widetilde{H}_{\text{AO}}$.

 @tbl-out-time shows the processing time of each estimator across different settings of $L$  and $n$.
<!-- <!-- # ```{r Simulated_comparative_time6, eval=FALSE, echo=FALSE} --> 
```{r Simulated_comparative_time6, echo=FALSE, message=FALSE}
set.seed(1234567890, kind = "Mersenne-Twister")
R <- 600
mu <- 1
B <- 40

sample.size <- c(25,49, 81, 121)
L_values <- c(5, 18)

bootstrap_estimators <- list(
  "$\\widetilde{H}_{\\text{C}}$" = bootstrap_correa_estimator,
  "$\\widetilde{H}_{\\text{AO}}$" = bootstrap_al_omari_1_estimator
)
if (!file.exists("./Data/results_data_time_sh.Rdata")) {#results_data_timenn
  set.seed(1234567890, kind = "Mersenne-Twister")

  all_TestStatistics <- list()
  execution_times <- list()

  for (estimator_name in names(bootstrap_estimators)) {
    TestStatistics <- list()

    for (L in L_values) {
      TestStat <- list()

      for (s in sample.size) {
        TestStat1 <- numeric(R)
        start_time <- Sys.time() # start time

        for (r in 1:R) {
          z <- gamma_sar_sample(L, mu, s)
          TestStat1[r] <- bootstrap_estimators[[estimator_name]](z, B) - (log(mean(z)) + (L - log(L) + lgamma(L) + (1 - L) * digamma(L)))
        }

        TestStatistics[[as.character(s)]] <- data.frame("SampleSize" = rep(s, R), "Test_Statistics" = TestStat1)

        end_time <- Sys.time() # end time
        execution_time <- end_time - start_time # total time
        execution_times[[paste(estimator_name, L, s)]] <- execution_time # save time
      }

      all_TestStatistics[[paste(estimator_name, L)]] <- TestStatistics
    }
  }

  save(all_TestStatistics, execution_times, file = "./Data/results_data_time_sh.Rdata")#
}

```
\renewcommand{\arraystretch}{1.1}
```{r Table_statistic_time_6,  echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
#| label: tbl-out-time
#| tbl-cap: 'Processing time for bootstrap-improved estimators $\widetilde{H}_{\text{C}}$ and $\widetilde{H}_{\text{AO}}$'

cat("\\setlength{\\tabcolsep}{10pt}\n")
load("./Data/results_data_time_sh.Rdata")
#fig.show="hide",
# eval=FALSE,
combined_results <- data.frame(
  Estimator   = character(),
  L           = numeric(),
  SampleSize  = numeric(),
  Time        = numeric()
)

# Organizar resultados
for (estimator_name in names(bootstrap_estimators)) {
  for (L in L_values) {
    for (sample_size in sample.size) {
      key <- paste(estimator_name, L, sample_size)
      if (!is.null(execution_times[[key]])) {
        combined_results <- rbind(combined_results, data.frame(
          Estimator   = estimator_name,
          L           = L,
          SampleSize  = sample_size,
          Time        = as.numeric(execution_times[[key]], units = "secs")
        ))
      }
    }
  }
}

# Filas para negrita (índices 9:16)
rows_to_bold <- 9:16

# Formato columnas de texto
combined_results$L          <- sprintf("$%d$", combined_results$L)
combined_results$SampleSize <- sprintf("$%d$", combined_results$SampleSize)
combined_results$Time       <- sapply(seq_along(combined_results$Time), function(i) {
  val <- combined_results$Time[i]
  if (i %in% rows_to_bold) {
    sprintf("$\\mathbf{\\phantom{-}%.2f}$", val)
  } else {
    sprintf("$\\phantom{-}%.2f$", val)
  }
})

# Renombrar columnas para LaTeX
colnames(combined_results) <- c("\\textbf{Estimator}", "$\\bm{L}$", "$\\bm{n}$", "\\textbf{Time} (s)")

# Generar tabla con kable (latex)
knitr::kable(
  combined_results,
#  caption    = "Processing time for bootstrap-improved estimators \\(\\widetilde{H}_{\\text{C}}\\) and \\(\\widetilde{H}_{\\text{AO}}\\).",
  format     = "latex",
  booktabs   = TRUE,
  align      = "cccc",
  escape     = FALSE,
  #label      = "table_time",
  #table.envir= "table",
 # position   = "H",
  linesep    = ""
) %>%
  row_spec(4,  extra_latex_after = "\\cline{3-4}") %>%
  row_spec(12, extra_latex_after = "\\cline{3-4}") %>%
  collapse_rows(latex_hline = "major", valign = "middle") %>%
  row_spec(0, align = "c") %>%
  kable_styling(latex_options = "scale_down", full_width = TRUE)

```

The results indicate that $\widetilde{H}_{\text{AO}}$ consistently achieves lower computational times compared to $\widetilde{H}_{\text{C}}$. This makes it the more efficient option, particularly relevant for large datasets, such as SAR images discussed in @sec-Chapter4. Consequently, $\widetilde{H}_{\text{AO}}$ is selected as the representative non-parametric Shannon entropy estimator in the simulations presented in the following sections.

## Optimal $\lambda$ Selection 
<!-- Finding the Optimal Value of $\lambda$ for the Rényi and Tsallis Entropy Estimators -->

We aim to determine the optimal order $\lambda$ for the nonparametric entropy estimators of Rényi ($\widetilde{R}_{\lambda}$) and Tsallis ($\widetilde{T}_{\lambda}$), using simulated samples of size $n = 49$ from $Z\sim\Gamma_{\text{SAR}}(5,  1)$ distribution.

To assess estimator performance, we conducted a Monte Carlo simulation with $R = 1000$ replications. For each value of $\lambda$, independent samples were generated, and entropy was estimated for each sample using the bootstrap bias-corrected version with $B = 100$ resamples  to reduce bias. The resulting estimates were compared to the theoretical entropy to compute the bias and MSE. This simulation-based procedure allows us to evaluate how the choice of $\lambda$ affects estimation accuracy, and whether the bootstrap improves performance.

For the Rényi estimator (see Fig. \ref{fig-optimal_order}), $\lambda = 0.9$ minimizes the MSE while keeping the bias low, achieving a favorable bias-variance trade-off. Although $\lambda = 0.85$ yields slightly lower bias, it produces a higher MSE, supporting the choice of $\lambda = 0.9$ as optimal.

Similarly, for the Tsallis estimator (Fig. \ref{fig-optimal_order-tsallis}), $\lambda = 0.85$ results in the lowest MSE with moderate bias, making it the most suitable choice.

```{r fig-optimal_order,  echo=FALSE, message=FALSE, warning=FALSE,  out.width="65%",  fig.pos="H", fig.cap="Bias and MSE as a function of $\\lambda$, for the Rényi entropy estimator, with $n = 49$, $L = 5$.", fig.width=5, fig.height=3}
#fig.show="hide", 
# eval=FALSE,  #

data <- data.frame(
  Lambda = c(0.9,  0.85, 0.99, 1.1, 1.5),
  Bias = c(0.00158,  -0.00250, 0.02077, 0.03751, 0.06512),
  MSE = c(0.01273,  0.01441, 0.01653, 0.01697, 0.01906)
)

data <- data[order(data$Lambda), ]

bias_plot <- ggplot(data, aes(x = Lambda, y = Bias)) +
  geom_line(color = "#00AFBB", linewidth = 1.0) +  
  geom_point(color = "#00AFBB", size = 3) +  
  geom_hline(yintercept = 0, color = "gray50", linetype = "dashed", linewidth = 0.5) +
  #annotate("text", x = max(data$Lambda) - 0.1, y = max(data$Bias), 
         #  label = "Bias", color = "#00AFBB", fontface = "italic", hjust = 1.0) +
  #scale_x_continuous(breaks = data$Lambda, labels = data$Lambda) +
  scale_x_continuous(breaks = data$Lambda, labels = data$Lambda, minor_breaks = NULL)+
  scale_y_continuous(minor_breaks = NULL)+
  labs(x = expression(lambda), y = "Bias") +
  theme_minimal() +
  theme(text = element_text(family = "serif"),
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 12))

mse_plot <- ggplot(data, aes(x = Lambda, y = MSE)) +
  geom_line(color = "#E69F00", linewidth = 1.0) +  
  geom_point(color = "#E69F00", size = 3) +
  #geom_hline(yintercept = 0, color = "gray50", linetype = "dashed", linewidth = 0.5) +
  #annotate("text", x = max(data$Lambda) - 0.1, y = max(data$MSE), 
           #label = "MSE", color = "#E69F00", fontface = "italic", hjust = 1.0) +
   #scale_x_continuous(breaks = data$Lambda, labels = data$Lambda) + 
  scale_x_continuous(breaks = data$Lambda, labels = data$Lambda, minor_breaks = NULL)+
  scale_y_continuous(minor_breaks = NULL)+
  labs(x = expression(lambda), y = "MSE") +
  theme_minimal() +
  theme(text = element_text(family = "serif"),
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 12))

grid.arrange(bias_plot, mse_plot, nrow = 2)

```

```{r fig-optimal_order-tsallis,  echo=FALSE, message=FALSE, warning=FALSE,  out.width="65%",  fig.pos="H", fig.cap="Bias and MSE as a function of $\\lambda$, for the Tsallis entropy estimator, with $n = 49$, $L = 5$.", fig.width=5, fig.height=3}
#fig.show="hide", 
# eval=FALSE,  #

data <- data.frame(
  Lambda = c(0.8,  0.85, 0.9, 0.95, 1.1, 1.5),
  Bias = c(-0.094,  0.00047, 0.009, 0.04,0.3, 0.6),
  MSE = c(0.0215,  0.022, 0.023, 0.026, 0.029, 0.031)
)

data <- data[order(data$Lambda), ]

bias_plot <- ggplot(data, aes(x = Lambda, y = Bias)) +
  geom_line(color = "#8E44AD", linewidth = 1.0) +  
  geom_point(color = "#8E44AD", size = 3) +  
  geom_hline(yintercept = 0, color = "gray50", linetype = "dashed", linewidth = 0.5) +
  #annotate("text", x = max(data$Lambda) - 0.1, y = max(data$Bias), 
         #  label = "Bias", color = "#00AFBB", fontface = "italic", hjust = 1.0) +
  #scale_x_continuous(breaks = data$Lambda, labels = data$Lambda) +
  scale_x_continuous(breaks = data$Lambda, labels = data$Lambda, minor_breaks = NULL)+
  scale_y_continuous(minor_breaks = NULL)+
  labs(x = expression(lambda), y = "Bias") +
  theme_minimal() +
  theme(text = element_text(family = "serif"),
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 12))

mse_plot <- ggplot(data, aes(x = Lambda, y = MSE)) +
  geom_line(color = "#27AE60", linewidth = 1.0) +  
  geom_point(color = "#27AE60", size = 3) +
  #geom_hline(yintercept = 0, color = "gray50", linetype = "dashed", linewidth = 0.5) +
  #annotate("text", x = max(data$Lambda) - 0.1, y = max(data$MSE), 
           #label = "MSE", color = "#E69F00", fontface = "italic", hjust = 1.0) +
   #scale_x_continuous(breaks = data$Lambda, labels = data$Lambda) + 
  scale_x_continuous(breaks = data$Lambda, labels = data$Lambda, minor_breaks = NULL)+
  scale_y_continuous(minor_breaks = NULL)+
  labs(x = expression(lambda), y = "MSE") +
  theme_minimal() +
  theme(text = element_text(family = "serif"),
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 12))

grid.arrange(bias_plot, mse_plot, nrow = 2)

```

<!-- ```{=latex} -->
<!-- \begin{figure}[hbt] -->
<!--   \centering -->
<!--   \includegraphics[width=0.42\textwidth]{Figures-R1/fig-optimal_order-1.pdf} -->
<!--   \vspace{0.5em} -->
<!--   \caption{Bias and MSE as a function of $\lambda$, with $n=49$, $L=5$.} -->
<!--   \label{fig-optimal_order} -->
<!-- \end{figure} -->
<!-- ``` -->
In the case of $L = 1$, the bootstrap procedure does not provide meaningful improvement and often leads to unstable estimates. Therefore, we omit bootstrap in this setting. Under these conditions, higher values of $\lambda$ yield better results: $\lambda = 3$ is preferred for the Rényi estimator, and $\lambda = 1.2$ for the Tsallis estimator.

### Bootstrap Bias Correction in Rényi and Tsallis Entropy Estimation

Building upon the previous Monte Carlo setup, we further assess the performance of the Rényi and Tsallis entropy estimators by analyzing their behavior across varying sample sizes $n \in \{9, 25, 49, 81, 121\}$. Using the optimal values of the order parameter ($\lambda = 0.9$ for Rényi and $\lambda = 0.85$ for Tsallis) already identified in the earlier analysis, we compute the bias and MSE for both the original and bootstrap-corrected versions of the estimators. As illustrated in Figures \ref{fig-bias_mse_rm1} and \ref{fig-bias_mse_Tsallis}, the bootstrap estimators consistently show improved performance, especially for small samples. In particular, at $n = 49$, both estimators exhibit low bias and MSE, confirming their robustness in moderate sample regimes and supporting their use in practical applications.

```{r Simulated_data_bias_m1, echo=FALSE, message=FALSE }
set.seed(1234567890, kind = "Mersenne-Twister")
#cache = TRUE, autodep = TRUE

file_name <- "./Data/results_renyi_m1.Rdata"

if (file.exists(file_name)) {

  load(file_name)
  message("Loaded existing results from results_renyi_m1.Rdata")
} else {
  # Parameters
  sample_sizes <- c(9, 25, 49, 81, 121)
  R <- 500
  B <- 100
  mu <- 1
  L <- 5
  alpha_values <- c(0.9)

  estimators <- list(
    "Renyi Estimator" = renyi_estimator,
    "Renyi Estimator Bootstrap" = bootstrap_renyi_estimator
  )

  results <- calculate_bias_mse_r2(sample_sizes, R, B, mu, L, alpha_values, estimators)

  save(results, file = file_name)
  message("Simulations completed and results saved.")
}
```

```{r fig-bias_mse_rm1,  echo=FALSE, message=FALSE, warning=FALSE,  out.width="70%",  fig.pos="H",  fig.cap="Bias and MSE of the Rényi entropy estimators for $\\Gamma_{\\text{SAR}}$, with $L=5$", fig.width=5.5, fig.height=4}

theme_set(theme_minimal() +
            theme(text = element_text(family = "serif"),
                  legend.position = "bottom"))
#fig.show="hide",
# eval=FALSE,  #
#  out.width="80%",
load("./Data/results_renyi_m1.Rdata")

alpha_values <- 0.9
estimators_to_plot <- c("Renyi Estimator", "Renyi Estimator Bootstrap")
latex_estimator_names <- c("Renyi Estimator" = expression("$\\widehat{italic(R)}_{\\lambda}$"),#
                           "Renyi Estimator Bootstrap" = expression("$\\widetilde{italic(R)}_{\\lambda}$"))
selected_estimators_latex <- latex_estimator_names[estimators_to_plot]

combined_plot_renyi <- generate_plot_renyi(results, alpha_values, selected_estimators_latex, ncol = 1, nrow = 1)

print(combined_plot_renyi)

```

```{r}
tsallis_entropy_gamma_sar <- function(mu, L, lambda) {
  (1 - exp((1 - lambda)*log(mu) +
             (lambda - 1)*log(L) +
             lgamma(lambda*(L - 1) + 1) -
             lambda*lgamma(L) -
             (lambda*(L - 1) + 1)*log(lambda))) / (lambda - 1)
}
# Función para calcular Bias y MSE para la entropía de Tsallis
calculate_bias_mse_tsallis <- function(sample_sizes, R, B, mu, L, lambda_values, estimators) {
  output <- data.frame(
    n = integer(0),
    Estimator = character(0),
    Lambda = numeric(0),
    Bias = numeric(0),
    MSE = numeric(0)
  )

  for (lambda in lambda_values) {
    true_entropy <- tsallis_entropy_gamma_sar(mu, L, lambda)

    for (ssize in sample_sizes) {
      samples <- generate_samples(ssize, R, mu, L)

      for (estimator_name in names(estimators)) {
        estimator <- estimators[[estimator_name]]
        v.entropy <- numeric(R)

        for (r in 1:R) {
          sample <- samples[[r]]

          if (grepl("Bootstrap", estimator_name)) {
            v.entropy[r] <- estimator(sample, B = B, lambda = lambda)
          } else {
            v.entropy[r] <- estimator(sample, lambda = lambda)
          }
        }

        mse <- mean((v.entropy - true_entropy)^2)
        bias <- mean(v.entropy) - true_entropy

        output <- rbind(output, data.frame(
          n = ssize,
          Estimator = estimator_name,
          Lambda = lambda,
          Bias = round(bias, 5),
          MSE = round(mse, 5)
        ))
      }
    }
  }

  return(output)
}

```

```{r Simulated_data_bias_tt, echo=FALSE, message=FALSE }
set.seed(1234567890, kind = "Mersenne-Twister")
#cache = TRUE, autodep = TRUE

file_name_tsallis <- "./Data/results_tsallis_jj.Rdata"

if (file.exists(file_name_tsallis)) {

  load(file_name_tsallis)
  message("Loaded existing results from results_tsallis_jj.Rdata")
